---
title: "RJDemetra: an R interface to JDemetra+"
subtitle: "NTTS, 13 March 2019"
author: "Alain Quartier-la-Tente and Anna Michalek"
division: "Insee, Seasonal Adjustment Centre of Excellence (AQLT) and European Central Bank (AM)"
logo: "img/SACElogo.jpg"
output:
    beamer_presentation:
        template: template.tex
        keep_tex: yes
        theme: TorinoTh
        slide_level: 3
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage[table]{xcolor}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{animate}
themeoptions: "coding=utf8,language=english"
classoption: 'usepdftitle=false,french'
fontsize: 10pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                        fig.path = "img/markdown-")
```

# RJDemetra

## Purpose and current status

### Purpose of the RJDemetra package

* Complete R package for Tramo-Seats and X13  
* Users: "pure R" package
    + Part of R routines, automatization
        - Batch processing
        - E.g.: direct vs indirect aggregates adjustment, dashboards
    + Usage of other R functions and packages
    
* JD+ functionality 
    + Modeling and seasonal adjustment 
    + Full specification
    
* Advanced graphical presentation: JD+ 



### Current status

- RegARIMA, TRAMO-SEATS and X-13-ARIMA:  
    + R package with documentation  
    + S3 classes with plot, summary, print methods
    + Possibility to add user-defined regressors but not user-defined calendar regressors

- Manipulate workspace (only TRAMO-SEATS and X-13-ARIMA):  
    + Import JD+ workspace to get: input raw series or SA model
    + Export R models created via RJDemetra
    

# Some examples
## RegARIMA examples
### RegARIMA examples (1/4)
\footnotesize

```{r}
library(RJDemetra);options(enable_print_style = FALSE)
myseries <- ipi_c_eu[,"FR"]
regarima_model <- regarima_def_x13(myseries, spec = "RG4c")
str(regarima_model, max.level = 1)
```

### RegARIMA examples (2/4)

\animategraphics[loop, autoplay, width=\linewidth]{4}{img/gif/regarima_print/img_}{1}{37}

### RegARIMA examples (3/4)

```{r, echo = FALSE}
def.par <- par(no.readonly = TRUE)
par(mai = c(0,0.82,0,0.42))
```

```{r, fig.height = 6 }
layout(matrix(1:6, 3, 2));plot(regarima_model, ask = FALSE)
```

### RegARIMA examples (3/4): extra plot

\footnotesize

```{r, echo = FALSE}
layout(1)
par(mai = c(0,0.82,0,0.42))
```

```{r, fig.height = 6}
plot(regarima_model, which = 7, dec_zoom = TRUE)
```



## Seasonal adjustment examples
### Seasonal adjustment examples (1/7)

A `SA` object is a `list()` of 5 elements:

1. `regarima`: the RegArima model
2. `decomposition`: decomposition variables ($\ne$ for TRAMO-SEATS and X-13-ARIMA)
3. `final`: time series main results
4. `diagnostics`: residuals tests, etc.
5. `user_defined`: other user_defined variables not exported by default (see `?user_defined_variables`)

\footnotesize
```{r}
x13_usr_spec <- x13_spec_def(spec=c("RSA5c"),usrdef.outliersEnabled = TRUE,
                             usrdef.outliersType = c("LS","AO"),
                             usrdef.outliersDate=c("2008-10-01","2002-01-01"),
                             usrdef.outliersCoef = c(36000,14000),
                             transform.function = "None")
x13_mod <- x13(myseries, x13_usr_spec)
ts_mod <- tramoseats_def(myseries, spec = "RSAfull")
```

### Seasonal adjustment examples (2/7)

\footnotesize

```{r}
x13_mod$decomposition
```

### Seasonal adjustment examples (3/7)

\footnotesize

```{r}
ts_mod$decomposition
```


### Seasonal adjustment examples (4/7)

```{r, fig.height=6}
plot(x13_mod$decomposition)
```

### Seasonal adjustment examples (5/7)

\footnotesize

```{r}
x13_mod$final
```

### Seasonal adjustment examples (6/7)

```{r,fig.height=6}
plot(x13_mod$final, first_date = 2012, type_chart = "sa-trend")
```


### Seasonal adjustment examples (7/7)

\footnotesize
```{r,fig.height=6}
x13_mod$diagnostics
```


## Manipulate workspaces


### Export a workspace

\footnotesize
```{r, echo= TRUE, message = FALSE}
wk <- new_workspace()
new_multiprocessing(wk, name = "MP-1")
add_sa_item(wk, multiprocessing = "MP-1",
            sa_obj = x13_mod, name =  "SA with X13 model 1 ")
add_sa_item(wk, multiprocessing =  "MP-1",
            sa_obj = ts_mod, name = "SA with TramoSeats model 1")
save_workspace(wk, "workspace.xml")
```

\includegraphics{img/workspace.png}

### Import a workspace (1/3)

\footnotesize

```{r}
wk <- load_workspace("workspace.xml")
get_ts(wk)
```

### Import a workspace (2/3)

\animategraphics[loop, autoplay, width=\linewidth]{7}{img/gif/import_model/}{1}{97}

### Import a workspace (3/3)
\footnotesize

```{r}
compute(wk) # Important to get the Sa model
models <- get_model(wk) # A progress bar is printed by default
# To extract only one model
mp <- get_object(wk, 1)
count(mp)
sa2 <- get_object(mp,2)
get_name(sa2)
mod <- get_model(wk, sa2)
```


## Installation and future development

### How to install the package?

The package is available on GitHub: https://github.com/jdemetra/rjdemetra

It has also it's own website: https://jdemetra.github.io/rjdemetra/

It package can be installed from CRAN:
```{r, eval = FALSE}
install.packages("RJDemetra")
```

Or from github (development version):

```{r, eval = FALSE}
devtools::install_github("jdemetra/rjdemetra")
```

\bcinfo To install it you need Java8:  in case you don't, install a portable version of Java8 and set the `JAVA_HOME` path.

## Future developments

### What's next? \bcpanchant (1/2)

Documentation:

- Vignette/article for the Journal of Statistical Software

- Guide to install the package with portable version of Java (when you don't have administrator rights)

- Cheat sheet

### What's next? \bcpanchant (2/2)

Package:

- Get only the Java object of a SA (to reduce computation/customize the output)

- Possibility to used user-defined calendar regressors (currently: only user-defined regressors)

- Function to "refresh" the model (JD+ 3.0.0)

### Why and how use RJDemetra?

\includegraphics[height = 1.5cm]{img/rjdqa_logo.png} A package for quality assessment for seasonal adjustment. It implements:

- Statistics Canada Dashboard (to provide a snapshot of an individual series at a point in time and points out some possible problems)

- Insee quality report matrix (used to help the analyst during production to prioritize the models to check)

$\rightarrow$ See the [Seasonal Adjustment handbook](https://ec.europa.eu/eurostat/web/products-manuals-and-guidelines/-/KS-GQ-18-001?inheritRedirect=true)

Available on github https://github.com/AQLT/rjdqa, still in development (only works for X13 models and no documentation yet)

Example of the dashboard:

```{r,fig.height=6.5, eval = FALSE}
library(rjdqa)
plot(sa_dashboard(x13_mod))
```

## Statistics Canada dashboard

### Example of the dashboard (1/2)
\footnotesize
```{r,fig.height=6.5, echo = FALSE}
library(rjdqa)
sa_db_data <- sa_dashboard(x13_mod)
plot(sa_db_data)
```

### Example of the dashboard (2/2)

1. **Recent History of Series**: plot of the raw series, the SA series and the trend for the most recent periods.  It is intended to identify trend direction, overall volatility and obvious outliers

2.  **Summary of Key Diagnostics**: key diagnostics as residual seasonality, recent and recurring outliers, moving seasonality, ARIMA model autocorrelation

3. **Estimated Patterns and Anticipated Movements**: estimated trading day, moving holiday and seasonal pattern (rescaled in additive decomposition to represent relative level)

4.  **Net Effect of Seasonal Adjustment**: movement in the raw series, compared to typical ranges centered around "neutral" value (when $SA_t = SA_{t-1}$)
